---
title: "WorldHappiness"
author: "Manuel Breve"
date: "5/15/2020"
output: html_document
---


```{r}
# starting from scratch
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE, include = TRUE, message = FALSE, warning = FALSE)
```


```{r message=FALSE}
# imports for the project
library(mice) # for imputing missing data
library(psych)
library(ggplot2) # for an improved toolset for plots
library(tableplot)
library(reshape2)
library(chemometrics) # for the mahalanobis distance
library(pracma)
library(PerformanceAnalytics) # to plot histograms
library(dplyr) # for an improved toolset
library(tibble) # rownames_to_column and viceversa


# clustering
library(FactoMineR)
library(factoextra)
library(plotrix)
library(gplots)
library(tableplot)

set.seed(42)
```

***

### Dataset
- [,1] Regional indicator:
- [,2] Ladder.score (RESPONSE VARIABLE) - national average response to the question of life evaluations. 
- [,3] Standard.error.of.ladder.score
- [,4] upperwhisker
- [,5] lowerwhisker
- [,6] Logged.GDP.per.capita
- [,7] Social.support - Having someone to count on in times of trouble.
- [,8] Healthy.life.expectancy - Health life expectancies at birth, based on data extracted from the WHO.
- [,9] Freedom.to.make.life.choices - Freedom to make life choices is the national average of responses to the GWP question.
- [,10] Generosity - Residual of regressing national average of response to the GWP question “Have you donated money to a charity in the past month?                      
- [,11] Perceptions.of.corruption -  national average of the survey responses to two questions in the GWP: “Is corruption widespread throughout the government or not” and “Is corruption widespread within businesses or not?”                 
- [,12] Ladder.score.in.Dystopia                  
- [,13] Explained.by..Log.GDP.per.capita         
- [,14] Explained.by..Social.support             
- [,15] Explained.by..Healthy.life.expectancy    
- [,16] Explained.by..Freedom.to.make.life.choices
- [,17] Explained.by..Generosity                 
- [,18] Explained.by..Perceptions.of.corruption  
- [,19] Dystopia...residual   

```{r}
path <- "data/WHR20_DataForFigure2.1.csv"
df <- read.csv(path, header = TRUE, sep = ",", dec=".", row.names=1)
summary(df)
```

```{r}
df <- df[c(1,6,7,8,9,10,11,19,2)]
names(df)
```

```{r}
df$Regional.indicator.num <- factor(df$Regional.indicator, levels=c('Central and Eastern Europe',
                                                                    'Commonwealth of Independent States',
                                                                    'East Asia',
                                                                    'Latin America and Caribbean',
                                                                    'Middle East and North Africa',
                                                                    'North America and ANZ',
                                                                    'South Asia',
                                                                    'Southeast Asia',
                                                                    'Sub-Saharan Africa',
                                                                    'Western Europe'
                                                                    )
                                , labels = 1:10)

df$Regional.indicator.num <- as.numeric(levels(df$Regional.indicator.num))[df$Regional.indicator.num]
df <- df %>% select(1,10, 2:9)
```

```{r}
chart.Correlation(df[2:10], histogram=TRUE)
```

***
### Outliers 
#### Univaritare Outliers 
```{r Boxplots}
gplot <- function(z){
  ggplot(data=df, aes(x=factor(''), y=df[,z])) + geom_boxplot(fill="azure3", outlier.color = "red", outlier.shape = 1) + theme(legend.position='none') + labs(title = paste(colnames(df)[z]), x="",y="")
}
p1 <- gplot(2)
p2 <- gplot(3)
p3 <- gplot(4)
p4 <- gplot(5)
p5 <- gplot(6)
p6 <- gplot(7)
p7 <- gplot(8)
p8 <- gplot(9)
p9 <- gplot(10)

gridExtra::grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,p9, nrow=3, ncol=3)
rm(p1,p2,p3,p4,p5,p6,p7,p8,p9)
```
An observation _x_ declared an *extreme outlier* if it lies outside of the interval $(Q1-3\cdot IQR, Q3+3\cdot IQR)$, where $IQR=Q3-Q1$ is called the Interquartile Range. 

An observation _x_ is declared a *mild outlier* if it lies outside of the interval  $(Q1-1.5\cdot IQR, Q3+1.5\cdot IQR)$

```{r}
uni.outlier <-function(z, mod=c("extreme","mild")){
  mod <- match.arg(mod)
  if (mod == "extreme") {
      extremelower <- as.numeric(quantile(df[,z],0.25)-(3*IQR(df[,z])))
      extremeupper <- as.numeric(quantile(df[,z],0.75)+(3*IQR(df[,z])))
      outliers <- df %>% rownames_to_column('city') %>%
        filter(df[,z]<extremelower | df[,z]>extremeupper) %>%
        column_to_rownames('city')
      return(outliers)
  } else if (mod == "mild") {
    extremelower <- as.numeric(quantile(df[,z],0.25)-(1.5*IQR(df[,z])))
    extremeupper <- as.numeric(quantile(df[,z],0.75)+(1.5*IQR(df[,z])))
    outliers <- df %>% rownames_to_column('city') %>%
      filter(df[,z]<extremelower | df[,z]>extremeupper) %>%
      column_to_rownames('city')
    return(outliers)
  } 
} 

gdp.uniout <- uni.outlier(3,mod = "extreme") # No outliers
gdp.uniout1 <- uni.outlier(3,mod = "mild")   # No outliers
support.uniout <- uni.outlier(4,mod = "extreme")  # No outliers
(support.uniout1 <- uni.outlier(4,mod = "mild")) 
health.uniout <- uni.outlier(5,mod = "extreme")   # No outliers
health.uniout1 <- uni.outlier(5,mod = "mild")     # No outliers
free.uniout <- uni.outlier(6,mod = "extreme")     # No outliers
(free.uniout1 <- uni.outlier(6,mod = "mild"))
gene.uniout <- uni.outlier(7,mod = "extreme")     # No outliers
(gene.uniout1 <- uni.outlier(7,mod = "mild"))
(corr.uniout <- uni.outlier(8,mod = "extreme"))
(corr.uniout1 <- uni.outlier(8,mod = "mild"))
dys.uniout <- uni.outlier(9,mod = "extreme")      # No outliers
(dys.uniout1 <- uni.outlier(9,mod = "mild")) 
ladder.uniout <- uni.outlier(10,mod = "extreme")   # No outliers
ladder.uniout1 <- uni.outlier(10,mod = "mild")    # No outliers


rm(gdp.uniout, gdp.uniout1, support.uniout, support.uniout1, health.uniout, health.uniout1, free.uniout, free.uniout1, gene.uniout, gene.uniout1, corr.uniout, corr.uniout1, dys.uniout, dys.uniout1, ladder.uniout, ladder.uniout1)
```


#### Multivariate Outliers
```{r}
outliers <- Moutlier(df, quantile = 0.975, plot = TRUE)
```

```{r}
set.seed(42)
outliers$cutoff

distances <- outliers$md
isOutlier <- (distances >= outliers$cutoff)

robust_distances <- outliers$rd
is_robust_Outlier <- (robust_distances >= outliers$cutoff)

table(isOutlier)
table(is_robust_Outlier)

```


```{r}
res.pca <- PCA(df, quali.sup = 1, quanti.sup= 9, scale.unit = TRUE)
```

```{r}
#Screeplot
fviz_screeplot(res.pca, addlabels = TRUE)
```

```{r}
#4 significant dimenions
significant_dimensions <- 5
coordinates <- res.pca$ind$coord[,1:significant_dimensions]
coordinates <- data.frame(coordinates)
hcpc <- HCPC(coordinates ,nb.clust = -1, consol = F, graph = FALSE)
barplot(hcpc$call$t$inert.gain[1:20])
```
```{r}
nc <- 6
final_hcpc <- HCPC (coordinates, nb.clust = nc, graph = TRUE)
```

```{r}
fviz_dend(final_hcpc, 
          cex = 0.7,                     # Label size
          palette = "jco",               # Color palette see ?ggpubr::ggpar
          rect = TRUE, rect_fill = TRUE, # Add rectangle around groups
          rect_border = "jco",           # Rectangle color
          #labels_track_height = 0.8      # Augment the room for labels
          )
```

```{r}
cut <- final_hcpc$data.clust$clust
table(cut)
catdes(cbind(as.factor(cut),df),1,proba = 0.0001)
```

```{r}
final_hcpc$desc.ind$para
```

```{r}
clusters <- cbind(res.pca$ind$coord[,1:2],cut,df)
clusters
```

